---
title: "R Notebook"
output: PDF File
---

Import data:
```{r}
library(readxl)
#subject 1
S001.45.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-45_2021-11-21_13-58-58_Test_VANTAN_VAN.txt", sep="")
S001.45.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-45_2021-11-21_13-59-39_Test_VANTAN_TAN.txt", sep="")
S001.sit.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-sitting_2021-11-21_13-56-54_Test_VANTAN_VAN.txt", sep="")
S001.sit.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-sitting_2021-11-21_13-57-34_Test_VANTAN_TAN.txt", sep="")
S001.stand.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-standing_2021-11-21_14-01-50_Test_VANTAN_VAN.txt", sep="")
S001.stand.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-standing_2021-11-21_14-02-20_Test_VANTAN_TAN.txt", sep="")
S001.supine.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-supine_2021-11-21_14-04-43_Test_VANTAN_VAN.txt", sep="")
S001.supine.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/001-supine_2021-11-21_14-05-26_Test_VANTAN_TAN.txt", sep="")
#vr
S001.45.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/001.sitting.45.csv")
S001.sit.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/001.sitting.csv")
S001.stand.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/001.standing.csv")
S001.supine.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/001.supine.csv")

#subject 2
#tablet
S002.45.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-45_2021-11-21_14-18-33_Test_VANTAN_VAN.txt", sep="")
S002.45.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-45_2021-11-21_14-19-09_Test_VANTAN_TAN.txt", sep="")
S002.sit.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-sitting_2021-11-21_14-20-23_Test_VANTAN_VAN.txt", sep="")
S002.sit.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-sitting_2021-11-21_14-20-58_Test_VANTAN_TAN.txt", sep="")
S002.stand.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-standing_2021-11-21_14-11-41_Test_VANTAN_VAN.txt", sep="")
S002.stand.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-standing_2021-11-21_14-12-24_Test_VANTAN_TAN.txt", sep="")
S002.supine.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-supine_2021-11-21_14-14-36_Test_VANTAN_VAN.txt", sep="")
S002.supine.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/002-supine_2021-11-21_14-15-17_Test_VANTAN_TAN.txt", sep="")
#vr
S002.45.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/002.sitting.45.csv")
S002.sit.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/002.sitting.csv")
S002.stand.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/002.standing.csv")
S002.supine.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/002.supine.csv")

#subject 3
#tablet
S003.45.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-45_2021-11-21_13-29-47_Test_VANTAN_VAN.txt", sep="")
S003.45.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-45_2021-11-21_13-30-24_Test_VANTAN_TAN.txt", sep="")
S003.sit.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-sitting_2021-11-21_13-27-06_Test_VANTAN_VAN.txt", sep="")
S003.sit.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-sitting_2021-11-21_13-27-56_Test_VANTAN_TAN.txt", sep="")
S003.stand.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-standing_2021-11-21_13-33-46_Test_VANTAN_VAN.txt", sep="")
S003.stand.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-standing_2021-11-21_13-34-25_Test_VANTAN_TAN.txt", sep="")
S003.supine.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-supine_2021-11-21_13-39-45_Test_VANTAN_VAN.txt", sep="")
S003.supine.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/003-supine_2021-11-21_13-40-18_Test_VANTAN_TAN.txt", sep="")
#vr
S003.45.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/003.sitting.45.csv")
S003.sit.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/003.sitting.csv")
S003.stand.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/003.standing.csv")
S003.supine.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/003.supine.csv")

#subject 4
#tablet
S004.45.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-45_2021-11-21_13-53-54_Test_VANTAN_VAN.txt", sep="")
S004.45.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-45_2021-11-21_13-54-32_Test_VANTAN_TAN.txt", sep="")
S004.sit.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-sitting_2021-11-21_13-51-54_Test_VANTAN_VAN.txt", sep="")
S004.sit.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-sitting_2021-11-21_13-52-34_Test_VANTAN_TAN.txt", sep="")
S004.stand.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-standing_2021-11-21_13-48-33_Test_VANTAN_VAN.txt", sep="")
S004.stand.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-standing_2021-11-21_13-49-14_Test_VANTAN_TAN.txt", sep="")
S004.supine.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-supine_2021-11-21_13-44-55_Test_VANTAN_VAN.txt", sep="")
S004.supine.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/004-supine_2021-11-21_13-45-51_Test_VANTAN_TAN.txt", sep="")
#vr
S004.45.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/004.sitting.45.csv")
S004.sit.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/004.sitting.csv")
S004.stand.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/004.standing.csv")
S004.supine.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/004.supine.csv")

#subject 5
#tablet
S005.45.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-45_2021-11-21_16-29-28_Test_VANTAN_VAN.txt", sep="")
S005.45.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-45_2021-11-21_16-30-15_Test_VANTAN_TAN.txt", sep="")
S005.sit.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-sitting_2021-11-21_16-34-23_Test_VANTAN_VAN.txt", sep="")
S005.sit.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-sitting_2021-11-21_16-35-08_Test_VANTAN_TAN.txt", sep="")
S005.stand.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-standing_2021-11-21_16-26-18_Test_VANTAN_VAN.txt", sep="")
S005.stand.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-standing_2021-11-21_16-26-59_Test_VANTAN_TAN.txt", sep="")
S005.supine.VAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-supine_2021-11-21_16-37-08_Test_VANTAN_VAN.txt", sep="")
S005.supine.TAN.tablet<- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/005-supine_2021-11-21_16-37-56_Test_VANTAN_TAN.txt", sep="")
#vr
S005.45.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/005.sitting.45.csv")
S005.sit.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/005.sitting.csv")
S005.stand.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/005.standing.csv")
S005.supine.VR <- read.csv("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN VR/005.supine.csv")

#subject distance from tablet
Subject_arm_length <- read_excel("~/Documents/GitHub/hsl_2021/VANTAN-VR-tab/VANTAN Tablet/Subject arm length.xlsx")

```

Clean data:
```{r}

#have to drop trials after 15 in VR datasets 
#45 deg condition
S001.45.VR <- S001.45.VR[c(1:15,21:35),]
S002.45.VR <- S002.45.VR[c(1:15,36:50),]
S003.45.VR <- S003.45.VR[c(1:15,21:35),]
S004.45.VR <- S004.45.VR[c(1:15,21:35),]
S005.45.VR <- S005.45.VR[c(1:15,21:35),]
#sitting condition
S001.sit.VR <- S001.sit.VR[c(1:15,21:35),]
S002.sit.VR <- S002.sit.VR[c(1:15,36:50),]
S003.sit.VR <- S003.sit.VR[c(1:15,21:35),]
S004.sit.VR <- S004.sit.VR[c(1:15,21:35),]
S005.sit.VR <- S005.sit.VR[c(1:15,21:35),]
#standing condition
S001.stand.VR <- S001.stand.VR[c(1:15,21:35),]
S002.stand.VR <- S002.stand.VR[c(1:15,37:51),]
S003.stand.VR <- S003.stand.VR[c(1:15,21:35),]
S004.stand.VR <- S004.stand.VR[c(1:15,21:35),]
S005.stand.VR <- S005.stand.VR[c(1:15,21:35),]
#supine condition 
S001.supine.VR <- S001.supine.VR[c(1:15,21:35),]
S002.supine.VR <- S002.supine.VR[c(1:15,36:50),]
S003.supine.VR <- S003.supine.VR[c(1:15,21:35),]
S004.supine.VR <- S004.supine.VR[c(1:15,21:35),]
S005.supine.VR <- S005.supine.VR[c(1:15,21:35),]

#Specific Corrections
#Subject 005 reported skipping trials 1-3 TAN VR Standing 
S005.stand.VR <- S005.stand.VR[-c(16:18),]
#Subject 003 reported skipping trial 1 TAN VR Sitting
S003.sit.VR <- S003.sit.VR[-c(16),]
#Subject 004 reported skipping trials 12-15 in VR VAN and TAN supine  
S004.supine.VR <- S004.supine.VR[-c(12:15,27:30),]

```

Put into big matrix:
```{r}

#EDIT: TAB and VR matrix; 8 col; 1:4 is VAN, 5:8 is TAN in (sit,stand,supine,45), 5 row = subject#)

TAB = rbind(cbind(S001.sit.VAN.tablet$YEnd,S001.stand.VAN.tablet$YEnd,S001.supine.VAN.tablet$YEnd,S001.45.VAN.tablet$YEnd,
                      S001.sit.TAN.tablet$AngEnd,S001.stand.TAN.tablet$AngEnd,S001.supine.TAN.tablet$AngEnd,S001.45.TAN.tablet$AngEnd),
                cbind(S002.sit.VAN.tablet$YEnd,S002.stand.VAN.tablet$YEnd,S002.supine.VAN.tablet$YEnd,S002.45.VAN.tablet$YEnd,
                      S002.sit.TAN.tablet$AngEnd,S002.stand.TAN.tablet$AngEnd,S002.supine.TAN.tablet$AngEnd,S002.45.TAN.tablet$AngEnd),
                cbind(S003.sit.VAN.tablet$YEnd,S003.stand.VAN.tablet$YEnd,S003.supine.VAN.tablet$YEnd,S003.45.VAN.tablet$YEnd,
                      S003.sit.TAN.tablet$AngEnd,S003.stand.TAN.tablet$AngEnd,S003.supine.TAN.tablet$AngEnd,S003.45.TAN.tablet$AngEnd),
                cbind(S004.sit.VAN.tablet$YEnd,S004.stand.VAN.tablet$YEnd,S004.supine.VAN.tablet$YEnd,S004.45.VAN.tablet$YEnd,
                      S004.sit.TAN.tablet$AngEnd,S004.stand.TAN.tablet$AngEnd,S004.supine.TAN.tablet$AngEnd,S004.45.TAN.tablet$AngEnd),
                cbind(S005.sit.VAN.tablet$YEnd,S005.stand.VAN.tablet$YEnd,S005.supine.VAN.tablet$YEnd,S005.45.VAN.tablet$YEnd,
                      S005.sit.TAN.tablet$AngEnd,S005.stand.TAN.tablet$AngEnd,S005.supine.TAN.tablet$AngEnd,S005.45.TAN.tablet$AngEnd)
          )
subjects_tab = rbind(t(t(rep(1, 11))),t(t(rep(2, 11))),t(t(rep(3, 11))),t(t(rep(4, 11))),t(t(rep(5, 11))))
TAB = cbind(subjects_tab,TAB)
colnames(TAB) = c("subjectID","sit_VAN","stand_VAN","supine_VAN","45_VAN","sit_TAN","stand_TAN","supine_TAN","45_TAN")

VR = rbind(
              cbind(S001.sit.VR$Accuracy[S001.sit.VR$Test=='vertical'],S001.stand.VR$Accuracy[S001.stand.VR$Test=='vertical'],S001.supine.VR$Accuracy[S001.supine.VR$Test=='vertical'],S001.45.VR$Accuracy[S001.45.VR$Test=='vertical'],
              S001.sit.VR$Accuracy[S001.sit.VR$Test=='tangential'],S001.stand.VR$Accuracy[S001.stand.VR$Test=='tangential'],S001.supine.VR$Accuracy[S001.supine.VR$Test=='tangential'],S001.45.VR$Accuracy[S001.45.VR$Test=='tangential']),
                cbind(S002.sit.VR$Accuracy[S002.sit.VR$Test=='vertical'],S002.stand.VR$Accuracy[S002.stand.VR$Test=='vertical'],S002.supine.VR$Accuracy[S002.supine.VR$Test=='vertical'],S002.45.VR$Accuracy[S002.45.VR$Test=='vertical'],
              S002.sit.VR$Accuracy[S002.sit.VR$Test=='tangential'],S002.stand.VR$Accuracy[S002.stand.VR$Test=='tangential'],S002.supine.VR$Accuracy[S002.supine.VR$Test=='tangential'],S002.45.VR$Accuracy[S002.45.VR$Test=='tangential']),
                cbind(S003.sit.VR$Accuracy[S003.sit.VR$Test=='vertical'],S003.stand.VR$Accuracy[S003.stand.VR$Test=='vertical'],S003.supine.VR$Accuracy[S003.supine.VR$Test=='vertical'],S003.45.VR$Accuracy[S003.45.VR$Test=='vertical'],
              S003.sit.VR$Accuracy[S003.sit.VR$Test=='tangential'],S003.stand.VR$Accuracy[S003.stand.VR$Test=='tangential'],S003.supine.VR$Accuracy[S003.supine.VR$Test=='tangential'],S003.45.VR$Accuracy[S003.45.VR$Test=='tangential']),
                cbind(S004.sit.VR$Accuracy[S004.sit.VR$Test=='vertical'],S004.stand.VR$Accuracy[S004.stand.VR$Test=='vertical'],S004.supine.VR$Accuracy[S004.supine.VR$Test=='vertical'],S004.45.VR$Accuracy[S004.45.VR$Test=='vertical'],
              S004.sit.VR$Accuracy[S004.sit.VR$Test=='tangential'],S004.stand.VR$Accuracy[S004.stand.VR$Test=='tangential'],S004.supine.VR$Accuracy[S004.supine.VR$Test=='tangential'],S004.45.VR$Accuracy[S004.45.VR$Test=='tangential']),
                cbind(S005.sit.VR$Accuracy[S005.sit.VR$Test=='vertical'],S005.stand.VR$Accuracy[S005.stand.VR$Test=='vertical'],S005.supine.VR$Accuracy[S005.supine.VR$Test=='vertical'],S005.45.VR$Accuracy[S005.45.VR$Test=='vertical'],
              S005.sit.VR$Accuracy[S005.sit.VR$Test=='tangential'],S005.stand.VR$Accuracy[S005.stand.VR$Test=='tangential'],S005.supine.VR$Accuracy[S005.supine.VR$Test=='tangential'],S005.45.VR$Accuracy[S005.45.VR$Test=='tangential'])
              
              )
subjects_vr = rbind(t(t(rep(1, 15))),t(t(rep(2, 15))),t(t(rep(3, 15))),t(t(rep(4, 15))),t(t(rep(5, 15))))
VR = cbind(subjects_vr,VR)
colnames(VR) = c("subjectID","sit_VAN","stand_VAN","supine_VAN","45_VAN","sit_TAN","stand_TAN","supine_TAN","45_TAN")


#Make it normal bc idk why it doesn't work otherwise
TAB = data.frame(data.matrix(TAB))
VR= data.frame(data.matrix(VR))

```


Convert Values to Angles
```{r}
#need to convert VAN data from raw to actual deviation
#vertical ocular misalignment = 180◦/pi*tan^−1(((pm − ps)/(pp))/d)
#where pM is the subject-selected position of the moving line in pixels, pS is the position ofthe stationary line in pixels, d is the distance
#between the subject’s eyes and the tablet screen in inches, and pp
#is the tablet’s resolution in pixels per inch

# VR values are not comparable to tablet ? problem?

TAB[1:11,2:5] = 180/pi*atan((TAB[1:11,2:5]/218)/Subject_arm_length$`Arm Length`[1])
TAB[12:22,2:5] = 180/pi*atan((TAB[12:22,2:5]/218)/Subject_arm_length$`Arm Length`[2])
TAB[23:33,2:5] = 180/pi*atan((TAB[23:33,2:5]/218)/Subject_arm_length$`Arm Length`[3])
TAB[34:44,2:5] = 180/pi*atan((TAB[34:44,2:5]/218)/Subject_arm_length$`Arm Length`[4])
TAB[45:55,2:5] = 180/pi*atan((TAB[45:55,2:5]/218)/Subject_arm_length$`Arm Length`[5])

```


VAN & TAN means
```{r}

TAB_avg = data.frame(data.matrix(
                    rbind(colMeans(TAB[1:11,2:9][sapply(TAB[1:11,2:9], is.numeric)]),
                    colMeans(TAB[12:22,2:9][sapply(TAB[12:22,2:9], is.numeric)]),
                    colMeans(TAB[23:33,2:9][sapply(TAB[23:33,2:9], is.numeric)]),
                    colMeans(TAB[34:44,2:9][sapply(TAB[34:44,2:9], is.numeric)]),
                    colMeans(TAB[45:55,2:9][sapply(TAB[45:55,2:9], is.numeric)])
                    )))

TAB_avg = cbind(c(1,2,3,4,5),TAB_avg)

VR_avg = data.frame(data.matrix(
                  rbind(colMeans(VR[1:15,2:9][sapply(VR[1:15,2:9], is.numeric)]),
                    colMeans(VR[16:30,2:9][sapply(VR[16:30,2:9], is.numeric)]),
                    colMeans(VR[31:45,2:9][sapply(VR[31:45,2:9], is.numeric)]),
                    colMeans(VR[46:60,2:9][sapply(VR[46:60,2:9], is.numeric)]),
                    colMeans(VR[60:75,2:9][sapply(VR[60:75,2:9], is.numeric)])
                    )))
VR_avg = cbind(c(1,2,3,4,5),VR_avg)

```

First test: paired-t, subjects matters (includes shapiro-wilks normality test)
```{r}

#t-tests comparing each condition in each modality, by instrument
ttest.bycondition.list = list()
for (i in 1:8) 
  {
    if ( shapiro.test(TAB_avg[,i+1])$p.value < 0.05 || shapiro.test(VR_avg[,i+1])$p.value < 0.05)
    {
      wilk = wilcox.test(TAB_avg[,i+1], VR_avg[,i+1], paired=TRUE)
      ttest.bycondition.list[[i]] = wilk$p.value
    } 
    else {
      ttest = t.test(TAB_avg[,i+1], VR_avg[,i+1], paired=TRUE)
      ttest.bycondition.list[[i]] = ttest$p.value
    }
  }
ttest.bycondition = data.frame(rbind(ttest.bycondition.list))
colnames(ttest.bycondition) = colnames(TAB_avg[,2:9])
print.data.frame(ttest.bycondition)
```

Second test: paired-t, subjects don't matter (includes shapiro-wilks normality test)
```{r}
#t-tests comparing each condition in each modality, by instrument
#preserve all VR data and apply unpaired
#ignoring for now
ttest.bycondition.pop.list = list()
for (i in 1:8) 
  {
    if ( shapiro.test(TAB[,i+1])$p.value < 0.05 || shapiro.test(VR[,i+1])$p.value < 0.05)
    {
      wilk = wilcox.test(TAB[,i+1], VR[,i+1], paired=FALSE)
      ttest.bycondition.pop.list[[i]] = wilk$p.value
    } 
    else {
      ttest = t.test(TAB[,i+1], VR[,i+1], paired=FALSE)
      ttest.bycondition.pop.list[[i]] = ttest$p.value
    }
  }
ttest.bycondition.pop = data.frame(rbind(ttest.bycondition.pop.list))
colnames(ttest.bycondition.pop) = colnames(TAB[,2:9])


#cut VR data and apply paired
VR.cut = rbind(VR[1:11,],VR[16:26,],VR[31:41,],VR[46:56,],VR[61:71,])
ttest.bycondition.popcut.list = list()
for (i in 1:8) 
  {
    if ( shapiro.test(TAB[,i+1])$p.value < 0.05 || shapiro.test(VR.cut[,i+1])$p.value < 0.05)
    {
      wilk = wilcox.test(TAB[,i+1], VR.cut[,i+1], paired=TRUE)
      ttest.bycondition.popcut.list[[i]] = wilk$p.value
    } 
    else {
      ttest = t.test(TAB[,i+1], VR.cut[,i+1], paired=TRUE)
      ttest.bycondition.popcut.list[[i]] = ttest$p.value
    }
  }
ttest.bycondition.popcut = data.frame(rbind(ttest.bycondition.popcut.list))
colnames(ttest.bycondition.popcut) = colnames(TAB[,2:9])
print.data.frame(ttest.bycondition.popcut)

```

Third test: paired-t, subjects matter, conditions don't matter - any overall differences between VR and TAB? (with avg data)
```{r}
VR_avg.pop = data.frame(cbind(stack(VR_avg[2:5])[,1],stack(VR_avg[6:9])[,1]))
colnames(VR_avg.pop) = c("VAN","TAN")
TAB_avg.pop = data.frame(cbind(stack(TAB_avg[2:5])[,1],stack(TAB_avg[6:9])[,1]))
colnames(TAB_avg.pop) = c("VAN","TAN")


ttest.conditiondm.pop.list = list()
for (i in 1:2) 
  {
    if ( shapiro.test(TAB_avg.pop[,i])$p.value < 0.05 || shapiro.test(VR_avg.pop[,i])$p.value < 0.05)
    {
      wilk = wilcox.test(TAB_avg.pop[,i], VR_avg.pop[,i], paired=TRUE)
      ttest.conditiondm.pop.list[[i]] = wilk$p.value
    } 
    else {
      ttest = t.test(TAB_avg.pop[,i], VR_avg.pop[,i], paired=TRUE)
      ttest.conditiondm.pop.list[[i]] = ttest$p.value
    }
  }
ttest.conditiondm.pop = data.frame(rbind(ttest.conditiondm.pop.list))
colnames(ttest.conditiondm.pop) = c("VAN","TAN")

print.data.frame(ttest.conditiondm.pop)

```

Fourth test: paired & unpaired t-test, subjects don't matter, conditions don't matter - any overall differences between VR and TAB? (with all data)
```{r}
# with VR cut
VR.pop = data.frame(cbind(stack(VR.cut[2:5])[,1],stack(VR.cut[6:9])[,1]))
colnames(VR.pop) = c("VAN","TAN")
TAB.pop = data.frame(cbind(stack(TAB[2:5])[,1],stack(TAB[6:9])[,1]))
colnames(TAB.pop) = c("VAN","TAN")

ttest.conditiondm.popdm.list = list()
for (i in 1:2) 
  {
    if (shapiro.test(TAB.pop[,i])$p.value < 0.05 || shapiro.test(VR.pop[,i])$p.value < 0.05)
    {
      wilk = wilcox.test(TAB.pop[,i], VR.pop[,i], paired=TRUE)
      ttest.conditiondm.popdm.list[[i]] = wilk$p.value
    } 
    else {
      ttest = t.test(TAB.pop[,i], VR.pop[,i], paired=TRUE)
      ttest.conditiondm.popdm.list[[i]] = ttest$p.value
    }
  }
ttest.conditiondm.popdm = data.frame(rbind(ttest.conditiondm.popdm.list))
colnames(ttest.conditiondm.popdm) = c("VAN","TAN")
print.data.frame(ttest.conditiondm.popdm)

```


Fifth test: one-way repeated measures ANOVA - are conditions are different WITHIN modalities & instruments
```{r}

# for TAB/VR avg, figure out if conditions are different within modality
# make factors base on conditions
conditionsVAN = rbind(t(t(rep(colnames(TAB_avg)[2],5))),t(t(rep(colnames(TAB_avg)[3],5))),t(t(rep(colnames(TAB_avg)[4],5))),t(t(rep(colnames(TAB_avg)[5],5))))
conditionsTAN = rbind(t(t(rep(colnames(TAB_avg)[6],5))),t(t(rep(colnames(TAB_avg)[7],5))),t(t(rep(colnames(TAB_avg)[8],5))),t(t(rep(colnames(TAB_avg)[9],5))))

TAB_avg.aov.van = data.frame(cbind(conditionsVAN,stack(TAB_avg[,2:5])[,1]))
TAB_avg.aov.van$X1 = as.factor(TAB_avg.aov.van$X1)
colnames(TAB_avg.aov.van) = c("conditionsVAN","VANtab")
cond.VAN.tab.aov = aov(VANtab ~ conditionsVAN, data = TAB_avg.aov.van)
summary(cond.VAN.tab.aov)

TAB_avg.aov.tan = data.frame(cbind(conditionsTAN,stack(TAB_avg[,6:9])[,1]))
TAB_avg.aov.tan$X1 = as.factor(TAB_avg.aov.tan$X1)
colnames(TAB_avg.aov.tan) = c("conditionsTAN","TANtab")
cond.TAN.tab.aov = aov(TANtab ~ conditionsTAN, data = TAB_avg.aov.tan)
summary(cond.TAN.tab.aov)

VR_avg.aov.van = data.frame(cbind(conditionsVAN,stack(VR_avg[,2:5])[,1]))
VR_avg.aov.van$X1 = as.factor(VR_avg.aov.van$X1)
colnames(VR_avg.aov.van) = c("conditionsVAN","VANvr")
cond.VAN.vr.aov = aov(VANvr ~ conditionsVAN, data = VR_avg.aov.van)
summary(cond.VAN.vr.aov)

VR_avg.aov.tan = data.frame(cbind(conditionsTAN,stack(VR_avg[,6:9])[,1]))
VR_avg.aov.tan$X1 = as.factor(VR_avg.aov.tan$X1)
colnames(VR_avg.aov.tan) = c("conditionsTAN","TANvr")
cond.TAN.vr.aov = aov(TANvr ~ conditionsTAN, data = VR_avg.aov.tan)
summary(cond.TAN.vr.aov)

```

Sixth test: one-way repeated measures ANOVA - are there differences across the conditions, between modalities (with avg data)
```{r}

#organize into VAN and TAN, mix modalities
conditionsVAN.mod = rbind(t(t(rep("sitTABvan",5))),t(t(rep("standTABvan",5))),t(t(rep("supineTABvan",5))),t(t(rep("45TABvan",5))),t(t(rep("sitVRvan",5))),t(t(rep("standVRvan",5))),t(t(rep("supineVRvan",5))),t(t(rep("45VRvan",5))))
conditionsTAN.mod = rbind(t(t(rep("sitTABtan",5))),t(t(rep("standTABtan",5))),t(t(rep("supineTABtan",5))),t(t(rep("45TABtan",5))),t(t(rep("sitVRtan",5))),t(t(rep("standVRtan",5))),t(t(rep("supineVRtan",5))),t(t(rep("45VRtan",5))))

VAN_avg = data.frame(cbind(TAB_avg$sit_VAN,TAB_avg$stand_VAN,TAB_avg$supine_VAN,TAB_avg$X45_VAN,
                VR_avg$sit_VAN,VR_avg$stand_VAN,VR_avg$supine_VAN,VR_avg$X45_VAN))
VAN_avg = data.frame(cbind(conditionsVAN.mod,stack(VAN_avg)[,1]))
VAN_avg$X1 = as.factor(VAN_avg$X1)
colnames(VAN_avg) = c("conditionsVAN.mod","VAN.mod")
cond.VAN_avg.aov = aov(VAN.mod ~ conditionsVAN.mod, data = VAN_avg)
summary(cond.VAN_avg.aov)

TAN_avg = data.frame(cbind(TAB_avg$sit_TAN,TAB_avg$stand_TAN,TAB_avg$supine_TAN,TAB_avg$X45_TAN,
                VR_avg$sit_TAN,VR_avg$stand_TAN,VR_avg$supine_TAN,VR_avg$X45_TAN))
TAN_avg = data.frame(cbind(conditionsTAN.mod,stack(TAN_avg)[,1]))
TAN_avg$X1 = as.factor(TAN_avg$X1)
colnames(TAN_avg) = c("conditionsTAN.mod","TAN.mod")
cond.TAN_avg.aov = aov(TAN.mod ~ conditionsTAN.mod, data = TAN_avg)
summary(cond.TAN_avg.aov)


```
Seventh test: one-way repeated measures ANOVA - are there differences across the conditions, between modalities (all data)
```{r}

#organize into VAN and TAN, mix modalities
conditionsVANall.mod = rbind(t(t(rep("sitTABvan",55))),t(t(rep("standTABvan",55))),t(t(rep("supineTABvan",55))),t(t(rep("45TABvan",55))),t(t(rep("sitVRvan",55))),t(t(rep("standVRvan",55))),t(t(rep("supineVRvan",55))),t(t(rep("45VRvan",55))))
conditionsTANall.mod = rbind(t(t(rep("sitTABtan",55))),t(t(rep("standTABtan",55))),t(t(rep("supineTABtan",55))),t(t(rep("45TABtan",55))),t(t(rep("sitVRtan",55))),t(t(rep("standVRtan",55))),t(t(rep("supineVRtan",55))),t(t(rep("45VRtan",55))))

VAN_all = data.frame(cbind(TAB$sit_VAN,TAB$stand_VAN,TAB$supine_VAN,TAB$X45_VAN,
                VR.cut$sit_VAN,VR.cut$stand_VAN,VR.cut$supine_VAN,VR.cut$X45_VAN))
VAN_all = data.frame(cbind(conditionsVANall.mod,stack(VAN_all)[,1]))
VAN_all$X1 = as.factor(VAN_all$X1)
colnames(VAN_all) = c("conditionsVANall.mod","VAN_all.mod")
cond.VAN_all.aov = aov(VAN_all.mod ~ conditionsVANall.mod, data = VAN_all)
summary(cond.VAN_all.aov)

TAN_all = data.frame(cbind(TAB$sit_TAN,TAB$stand_TAN,TAB$supine_TAN,TAB$X45_TAN,
                VR.cut$sit_TAN,VR.cut$stand_TAN,VR.cut$supine_TAN,VR.cut$X45_TAN))
TAN_all = data.frame(cbind(conditionsTANall.mod,stack(TAN_all)[,1]))
TAN_all$X1 = as.factor(TAN_all$X1)
colnames(TAN_all) = c("conditionsTANall.mod","TAN_all.mod")
cond.TAN_all.aov = aov(TAN_all.mod ~ conditionsTANall.mod, data = TAN_all)
summary(cond.TAN_all.aov)


```


Eighth test: two way repeated measures ANOVA - is alignment affected by condition and modality (with avg data)
```{r}
# set up array using 2 factors: conditions (4) and modality (2)
#conditionVAN.nomod =  rbind(t(t(rep("sit_van",5))),t(t(rep("stand_van",5))),t(t(rep("supine_van",5))),t(t(rep("45_van",5))),t(t(rep("sit_van",5))),t(t(rep("stand_van",5))),t(t(rep("supine_van",5))),t(t(rep("45_van",5))))

conditionVAN.nomod =  rep(rep(c("sit_van","stand_van","supine_van","45_van"),each = 5),2)

conditionTAN.nomod =  rep(rep(c("sit_tan","stand_tan","supine_tan","45_tan"),each = 5),2)

modVAN = rbind(t(t(rep("TABvan",20))),t(t(rep("VRvan",20))))
VAN_avg.2aov = data.frame(cbind(modVAN,conditionVAN.nomod,VAN_avg$VAN.mod))
colnames(VAN_avg.2aov) = c("mod","cond","VAN")
VAN_avg.2aov$mod = as.factor(VAN_avg.2aov$mod)
VAN_avg.2aov$cond = as.factor(VAN_avg.2aov$cond)
VAN_avg.2aov.aov = aov(VAN ~ cond * mod,data = VAN_avg.2aov)
summary(VAN_avg.2aov.aov)

modTAN = rbind(t(t(rep("TABtan",20))),t(t(rep("VRtan",20))))
TAN_avg.2aov = data.frame(cbind(modTAN,conditionTAN.nomod,TAN_avg$TAN.mod))
colnames(TAN_avg.2aov) = c("mod","cond","TAN")
TAN_avg.2aov$mod = as.factor(TAN_avg.2aov$mod)
TAN_avg.2aov$cond = as.factor(TAN_avg.2aov$cond)
TAN_avg.2aov.aov = aov(TAN ~ cond * mod,data = TAN_avg.2aov)
summary(TAN_avg.2aov.aov)

```
Ninth test: two way repeated measures ANOVA - is alignment affected by condition and modality (with all data)
```{r}
# set up array using 2 factors: conditions (4) and modality (2)
#conditionVAN.nomod =  rbind(t(t(rep("sit_van",5))),t(t(rep("stand_van",5))),t(t(rep("supine_van",5))),t(t(rep("45_van",5))),t(t(rep("sit_van",5))),t(t(rep("stand_van",5))),t(t(rep("supine_van",5))),t(t(rep("45_van",5))))

conditionVAN_all.nomod =  rep(rep(c("sit_van","stand_van","supine_van","45_van"),each = 55),2)

conditionTAN_all.nomod =  rep(rep(c("sit_tan","stand_tan","supine_tan","45_tan"),each = 55),2)

modVAN.all = rbind(t(t(rep("TABvan",55))),t(t(rep("VRvan",55))))
VAN_all.2aov = data.frame(cbind(modVAN.all,conditionVAN_all.nomod,VAN_all$VAN_all.mod))
colnames(VAN_all.2aov) = c("mod_all","cond_all","VAN_all")
VAN_all.2aov$mod_all = as.factor(VAN_all.2aov$mod_all)
VAN_all.2aov$cond_all = as.factor(VAN_all.2aov$cond_all)
VAN_all.2aov.aov = aov(VAN_all ~ cond_all * mod_all,data = VAN_all.2aov)
summary(VAN_all.2aov.aov)

modTAN.all = rbind(t(t(rep("TABtan",55))),t(t(rep("VRtan",55))))
TAN_all.2aov = data.frame(cbind(modTAN.all,conditionTAN_all.nomod,TAN_all$TAN_all.mod))
colnames(TAN_all.2aov) = c("mod_all","cond_all","TAN_all")
TAN_all.2aov$mod_all = as.factor(TAN_all.2aov$mod_all)
TAN_all.2aov$cond_all = as.factor(TAN_all.2aov$cond_all)
TAN_all.2aov.aov = aov(TAN_all ~ cond_all * mod_all,data = TAN_all.2aov)
summary(TAN_avg.2aov.aov)

```
Tenth test: some wacky clustering stuff - VAN
```{r}

VAN_clust = data.frame(cbind(c(TAB$sit_VAN,VR.cut$sit_VAN),c(TAB$stand_VAN,VR.cut$stand_VAN),c(TAB$supine_VAN,VR.cut$supine_VAN),c(TAB$X45_VAN,VR.cut$X45_VAN)))
VAN_clust = data.frame(cbind(modVAN.all,VAN_clust))
colnames(VAN_clust) = c("modality","sit","stand","supine","45deg")


library(mclust);
#rownames(VAN_clust) = VAN_clust[,1];
mc_van = Mclust(VAN_clust);
mc_van$classification;
mc_van$G
km  = kmeans(VAN_clust[,2:5], centers = 8);
km$cluster; # what does this mean @?@?@????!?!

#plot(VAN_clust)
i = 2;j = 3; 
plot(VAN_clust[,i], VAN_clust[,j], xlab='sitVAN', ylab='standVAN',main="Sit vs Stand VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}

i = 2;j = 4;
plot(VAN_clust[,i], VAN_clust[,j], xlab='sitVAN', ylab='supineVAN',main="Sit vs Supine VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}

i = 2;j = 5; 
plot(VAN_clust[,i], VAN_clust[,j], xlab='sitVAN', ylab='45degVAN',main="Sit vs 45deg HT VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}

i = 3;j = 4; 
plot(VAN_clust[,i], VAN_clust[,j], xlab='standVAN', ylab='supineVAN',main="Stand vs Supine VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}

i = 3;j = 5; 
plot(VAN_clust[,i], VAN_clust[,j], xlab='standVAN', ylab='45degVAN',main="Stand vs 45deg HT VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}

i = 4;j = 5; 
plot(VAN_clust[,i], VAN_clust[,j], xlab='supineVAN', ylab='45degVAN',main="Supine vs 45deg HT VAN" ,type="n")
for(s in 1:nrow(VAN_clust)) {
  text(VAN_clust[s,i], VAN_clust[s,j], VAN_clust[s,1], col=km$cluster[s])
}


```


Eleventh test: some wacky clustering stuff - TAN
```{r}

TAN_clust = data.frame(cbind(c(TAB$sit_TAN,VR.cut$sit_TAN),c(TAB$stand_TAN,VR.cut$stand_TAN),c(TAB$supine_TAN,VR.cut$supine_TAN),c(TAB$X45_TAN,VR.cut$X45_TAN)))
TAN_clust = data.frame(cbind(modTAN.all,TAN_clust))
colnames(TAN_clust) = c("modality","sit","stand","supine","45deg")

library(mclust);
#rownames(VAN_clust) = VAN_clust[,1];
mc_tan = Mclust(TAN_clust);
mc_tan$classification;
mc_tan$G;
km  = kmeans(TAN_clust[,2:5], centers = 1);
km$cluster; # what does this mean @?@?@????!?!

#plot(TAN_clust)
i = 2;j = 3; # sit v stand
plot(TAN_clust[,i], TAN_clust[,j], xlab='sitTAN', ylab='standTAN',main="Sit vs Stand TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

i = 2;j = 4;
plot(TAN_clust[,i], TAN_clust[,j], xlab='sitTAN', ylab='supineTAN',main="Sit vs Supine TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

i = 2;j = 5; 
plot(TAN_clust[,i], TAN_clust[,j], xlab='sitTAN', ylab='45degTAN',main="Sit vs 45deg HT TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

i = 3;j = 4; 
plot(TAN_clust[,i], TAN_clust[,j], xlab='standTAN', ylab='supineTAN',main="Stand vs Supine TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

i = 3;j = 5; 
plot(TAN_clust[,i], TAN_clust[,j], xlab='standTAN', ylab='45degTAN',main="Stand vs 45deg HT TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

i = 4;j = 5; 
plot(TAN_clust[,i], TAN_clust[,j], xlab='supineTAN', ylab='45degTAN',main="Supine vs 45deg HT TAN" ,type="n")
for(s in 1:nrow(TAN_clust)) {
  text(TAN_clust[s,i], TAN_clust[s,j], TAN_clust[s,1], col=km$cluster[s])
}

```

